Slavnov products and KP tau functions

### 2025-05-08 draft.tex
finish calculations, start conclusions

\section{Tau functions}

@@

\section{Homogeneous limit}

@@

\section{Examples} Let us now consider the cases \(n=2, 3\) (the case \(n=1\) is trivial)
to gain some understanding of the problem.

@@

\section{Baker-Akhiezer function}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Baker-Akhiezer function}

@@

\subsubsection{BA in \(z\)-coordinates}
Using the expression~(\ref{eq:tau-function-miwa}), we can write the
Baker-Akhiezer function in terms of the \(z\)-coordinates.
Let us first write
\begin{equation}
  t_p - \lambda^{-p}/p = \frac{1}{p} \sum_{j=1}^n z_j^ p - \lambda^{-p}/p \; , 
\end{equation}
therefore
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t} - [\lambda^{-1}], w^{-1})} & = \exp\left[ \sum_{p=1}^\infty \left(t_p  - \frac{\lambda^{-p}}{p}\right) w^{-p} \right]\\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ - \sum_{p=1}^\infty \frac{\lambda^{-p}}{p} w^{-p} \right] \\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ \ln ( 1 - 1 / (\lambda w)) \right]
      = \left( 1 - \frac{1}{\lambda w}\right)  \prod_{j=1}^n \frac{1}{1 -  z_j /w} \\
    & = \frac{w^{n-1}}{\lambda } \left( \lambda w - 1 \right)  \prod_{j=1}^n \frac{1}{w -  z_j} \\
  \end{split}
\end{equation}

@@

Putting all these facts together, we write the Baker-Akhiezer function
as 
\begin{equation}
  \psi(\lambda; \bm{z}, \bm{v}) =
  e^{\xi(\bm{t}, \lambda)}  \left[ 1 + 
  \sum_{p=1}^n (-1)^p \lambda^{-p} \sigma_p^{(n)}(\bm{z}^{-1})
    + \frac{1}{\bm{\tau}(\bm{z}, \bm{v})}
    \det \widehat{\mathcal{K}}_{[\bm{F} \to n-row]}
  \right]\; .
\end{equation}
Next section we find a better expression for the Baker-Akhiezer
function.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fredholm determinant}

@@

We finally conclude that the Bakher-Akhiezer function can be written as
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \det \left( \mathbb{1} + 
    \frac{1}{\lambda} \mathcal{M}
  \right)\; , 
\end{equation}
where \(\mathcal{M} = \mathcal{K}^{-1} \check{\mathcal{K}}\) is a
finite dimensional matrix. Let us now use the elementary expression
\begin{subequations}
\begin{equation}
  \begin{split}
    \det \left( \mathbb{1} + \frac{1}{\lambda} \mathcal{M}\right)
    & = \exp \left[ \Tr \log \left( \mathbb{1} + \frac{1}{\lambda} \mathcal{M} \right) \right]\\
    & = \prod_{l=1}^n \left( 1 + \frac{\mu_l}{\lambda} \right) \; ,
  \end{split}
\end{equation}
\end{subequations}
where \(\bm{\mu} = \{\mu_l\}_{l=1}^n\) are the eigenvalues of the
matrix \(\mathcal{M}\), and these obviously depend on the Bethe roots
\(\bm{v}\) and parameters \(\bm{t}\). Therefore, we have
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \left(1 + \sum_{k=1}^n \frac{\xi_k(\bm{t})}{\lambda^k} 
  \right)\; . 
\end{equation}
where \(\xi_k(\bm{t}) = \sigma_k^{(n)}(\mu_1, \dots, \mu_n)\) and \(\mu_j = \mu_j(\bm{t}, \bm{v})\).

We can also write 
\begin{equation}
  \begin{split}
    \det \left( \mathbb{1} + \frac{1}{\lambda} \mathcal{M}\right)
    & = \prod_{l=1}^n \left( 1 + \frac{\mu_l}{\lambda} \right) \\ 
  \end{split}
\end{equation}

In the limit \(n \to \infty\) with with \(n/L \to 0\), we have
that the Baker-Akhiezer function can be written as
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \left( 1 + \sum_{k=1}^n \frac{(-1)^k}{\lambda^k} \sigma_k(\{\bm{p}\})
  \right)\; . 
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discussion}

It seems that we can consider a basis of functions \(\mathcal{Y}_j(z)\) for the Baker-Akhiezer
functions. I think that this object might say something interesting. In particular, we can
explicitly build solutions of the KP equation. In fact, the functions \(\mathcal{Y}_j\)
above have poles at the Bethe roots, it is interesting. 
\begin{itemize}
\item[\faCheck] What is the pole and root structure of the tau and Baker-Akhiezer (BA) functions?
  In particular, what happens at the Bethe roots?
\item[\faCheck] Basis tau functions
\item[\faCheck] Write the cases n=2 and n=3
\item[\faCheck] What the general case for all \(z_k\) approaching to a single
  value \(w\). 
\item[\faCheck] Write these tau functions in terms of Miwa coordinates
\item[\faCheck] I need to calculate the Baker-Akhiezer function for this model using the Japanese formula
  in terms of tau functions.
\item[\faCheck]  Write some explicit expressions for \(N=2\) and \(N=3\). 
\item Discuss the Toda hierarchy case. 
\item See the Limit \(L \to \infty\) and \(N \to \infty\) but sufficiently slower than \(L\). 
  I think that the matrix \(\mathcal{T}\) is a trace class operator and we write the BA function
  as a Fredholm determinant. 
\item Better investigate these functions in Miwa coordinates. See Fredholm determinant. 
  I think that this BA function will be similar to the soliton-like solution discussed in section 3.6.2 (page 41) of~\cite{Zabrodin2018}. 
  My functions seem to have a pole of order \(N\) at infinity. And \(N\) zeros at the Bethe roots. 
\item It turns out that the Baker-Akhiezer function is the appropriate object to discuss
  the Bethe roots. 
  \item I think that these tau functions come from algebro geometric tau functions. 
  \item I can try to study the dynamics closer to the Bethe roots, I think that the
    \(\Omega\)'s will have some interesting expansions, some of them might vanish. 
  \item This might become a Riemann Hilbert problem
  \item Are these tau functions associated to the Calogero-Moser model??
  \item Elliptic case - see slavnov, zabrodin's paper
\end{itemize}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tau functions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Homogeneous limit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let us now consider the cases \(n=2, 3\) (the case \(n=1\) is trivial)
to gain some understanding of the problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Baker-Akhiezer function}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Baker-Akhiezer  in \(z\)-coordinates}

@@

Using the expression~(\ref{eq:tau-function-miwa}), we can write the
Baker-Akhiezer function in terms of the \(z\)-coordinates.
Let us first write
\begin{equation}
  t_p - \lambda^{-p}/p = \frac{1}{p} \sum_{j=1}^n z_j^ p - \lambda^{-p}/p \; , 
\end{equation}
therefore
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t} - [\lambda^{-1}], w^{-1})} & = \exp\left[ \sum_{p=1}^\infty \left(t_p  - \frac{\lambda^{-p}}{p}\right) w^{-p} \right]\\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ - \sum_{p=1}^\infty \frac{\lambda^{-p}}{p} w^{-p} \right] \\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ \ln ( 1 - 1 / (\lambda w)) \right]
      = \left( 1 - \frac{1}{\lambda w}\right)  \prod_{j=1}^n \frac{1}{1 -  z_j /w} \\
    & = \frac{w^{n-1}}{\lambda } \left( \lambda w - 1 \right)  \prod_{j=1}^n \frac{1}{w -  z_j} \\
  \end{split}
\end{equation}

@@

Putting all these facts together, we write the Baker-Akhiezer function
as 
\begin{equation}
  \psi(\lambda; \bm{z}, \bm{v}) =
  e^{\xi(\bm{t}, \lambda)}  \left[ 1 + 
  \sum_{p=1}^n (-1)^p \lambda^{-p} \sigma_p^{(n)}(\bm{z}^{-1})
    + \frac{1}{\bm{\tau}(\bm{z}, \bm{v})}
    \det \widehat{\mathcal{K}}_{[\bm{F} \to n-row]}
  \right]\; .
\end{equation}
Next section we find a better expression for the Baker-Akhiezer
function, but from this expression we can see that all the poles
are at the point \(\lambda =0\).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Baker-Akhiezer in \(t\)-coordinates}

@@

We finally conclude that the Bakher-Akhiezer function can be written as
\begin{equation}
\label{eq:ba-fredholm}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \det \left( \mathbb{1} + 
    \frac{1}{\lambda} \mathcal{M}
  \right)\; , 
\end{equation}
where \(\mathcal{M} = \mathcal{K}^{-1} \check{\mathcal{K}}\) is a
finite dimensional matrix. Let us now use the elementary expression
\begin{subequations}
\begin{equation}
  \begin{split}
    \det \left( \mathbb{1} + \frac{1}{\lambda} \mathcal{M}\right)
    & = \exp \left[ \Tr \log \left( \mathbb{1} + \frac{1}{\lambda} \mathcal{M} \right) \right]\\
    & = \prod_{l=1}^n \left( 1 + \frac{\mu_l}{\lambda} \right) \; ,
  \end{split}
\end{equation}
\end{subequations}
where \(\bm{\mu} = \{\mu_l\}_{l=1}^n\) are the eigenvalues of the
matrix \(\mathcal{M}\), and these obviously depend on the Bethe roots
\(\bm{v}\) and parameters \(\bm{t}\). Therefore, we have
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \left(1 + \sum_{k=1}^n \frac{\xi_k(\bm{t})}{\lambda^k} 
  \right)\; . 
\end{equation}
where \(\xi_k(\bm{t}) = \sigma_k^{(n)}(\mu_1, \dots, \mu_n)\) and \(\mu_j = \mu_j(\bm{t}, \bm{v})\).

We can also write 
\begin{equation}
  \begin{split}
    \det \left( \mathbb{1} + \frac{1}{\lambda} \mathcal{M}\right)
    & = \exp\left[ \sum_{l=1}^n \ln \left( 1 + \frac{\mu_l}{\lambda} \right)\right]
    = \exp\left[ \sum_{l=1}^n \sum_{k=1}^\infty \frac{(-1)^{k-1}}{k} \left(\frac{\mu_l}{\lambda} \right)^k\right] \\ 
    & = \exp\left[ \sum_{k=1}^\infty \frac{(-1)^{k-1}}{\lambda^k} \gamma_k\right] \; ,
  \end{split}
\end{equation}
where we define the Miwa coordinates \(\gamma_k = \frac{1}{k}\sum_{l=1}^n \mu_l^k \). Consequently
\begin{equation}
    \det \left( \mathbb{1} + \frac{1}{\lambda} \mathcal{M}\right)
     = 1 + \sum_{k\geq 1}\frac{1}{\lambda^k} \sigma_k(\bm{\gamma})\; .
\end{equation}
Putting all these facts together, we can write the functions
\(\xi_k(\bm{t})\) in terms of \(\bm{\gamma}\).  Observe that now the
parameter \(n\) is hidden in the definition of the coordinates
\(\gamma_k\).  One advantgage of this expression is that now we seem
ot have a better setting to consider the thermodynamic limit \(n \to
\infty\) with with \(n/L \to 0\). In fact, the
expression~(\ref{eq:ba-fredholm}) suggests that in this limit, the
Baker-Akhiezer function in the thermodynamics limit can be written as
a Fredholm determinant. We are currently exploring this question and
hope to report new results soon.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this paper we have discussed some properties of the Slavnov
products and their relation with the tau functions of the KP
hierarchy.

An important initial observation is that the general form of this tau
function is written in terms of some alternant matrix, and this form
is known since (cite papers on the matrix models and so on). But the
functions associated to the current paper is drastically different
from those papers, and do not expand a point in the
Grassmannian. Therefore, it is necessary to better understand some of
these properties. In this paper we start exploring these results.

....

We have also started our exploration of the Baker-Akhiezer function since it seems to
be the relevant object of our investigation.

Soem future interesting directions. We might want to calculate the
solutions of the KP equation to see what they really mean. Perhaps
consider analitic and numerical calculations, since the calculation of
the determinants become complicated even for small values of \(n\).

More interestingly, one may try to consider the thermodynamic limit of
the Slavnov product and explore what it means from the viewpoint of
the Baker-Akhiezer function.  See the Limit \(L \to \infty\) and \(N
\to \infty\) but sufficiently slower than \(L\).  I think that the
matrix \(\mathcal{M}\) is a trace class operator and we write the BA
function as a Fredholm determinant.

I think that this BA function will be similar to the soliton-like
solution discussed in section 3.6.2 (page 41) of~\cite{Zabrodin2018}.
My functions seem to have a pole of order \(N\) at infinity. And \(N\)
zeros at the Bethe roots.

Finally, we can explore the elliptic case and see if one can see if
there exists any similar relation on the elliptic case.  Elliptic case
- see slavnov, zabrodin's paper
